\ZpL: a p-adic precision package

### 2018-02-04 ZpL/ZpL.tex

If provably correct results are desired, interval arithmetic provides
an alternative to floating point.  Instead of just tracking an approximation to the answer,
the package also tracks a radius within which the true result lies.
This method is commonly used for $p$-adic computations since
the ultrametric property of $p$-adic fields frequently keeps the radius small.
Computations remain fairly efficient with this approach, but numerical instability
can still lead to dramatic losses in precision, (see Section \ref{sec:demo} for many examples).
Tracking the precision of multiple variables concurrently, the set of possible
true values associated to an inexact value takes the form of an ellipsoid
with axes parallel to the coordinate axes.

@@

\caption{The distribution of $\cohgt_t(\ttw)$}
\label{fig:deletion}
\end{figure}
%
The histogram of Figure~\ref{fig:deletion} shows 
the distribution of $\cohgt_t(\ttw)$ while executing the Euclidean
algorithm (naive implementation) with two polynomials of degree $7$
as input. The bias is evident: most of the time $\cohgt_t(\ttw) \leq
1$.

@@

%\begin{itemize}
%\item In which case should we use this package? Mainly for exploration
%\item example of $p$-adic differential equations (especially for $p=2$)
%\end{itemize}

%\subsection{Exploration}

>>>>

If provably correct results are desired, interval arithmetic provides
an alternative to floating point.  Instead of just tracking an approximation to the answer,
the package also tracks a radius within which the true result lies.
This method is commonly used for $p$-adic computations since
the ultrametric property of $p$-adic fields frequently keeps the radius small.
Computations remain fairly efficient with this approach, but numerical instability
can still lead to dramatic losses in precision (see Section \ref{sec:demo} for many examples).
Tracking the precision of multiple variables concurrently, the set of possible
true values associated to an inexact value takes the form of an ellipsoid
with axes parallel to the coordinate axes.

@@

\caption{The distribution of $\cohgt_t(\ttw)$}
\label{fig:deletion}
\end{figure}
%
The histogram of Figure~\ref{fig:deletion} shows 
the distribution of $\cohgt_t(\ttw)$ while executing the Euclidean
algorithm (naive implementation) with two polynomials of degree $7$
as input. The bias is evident: most of the time $\cohgt_t(\ttw) \leq
1$.
%For example, while executing the Euclidean algorithm (naive 
%implementation) with two polynomials of degree $7$, most of the 
%time $\cohgt_t(\ttw) \leq 1$.

@@

The package \ZpL provides powerful tools (based on Automatic 
Differentiation) to track precision in the $p$-adic setting. In many 
concrete situations, it heavily outdoes usual interval arithmetics 
which is currently widely used by Computer Algebra softwares (see
\S \ref{sec:demo} for many examples).
The impact of complexity is controlled but nevertheless non-negligible 
(see \S \ref{ssec:complexity}). For this reason, it is not really 
reasonable to believe that a fast algorithm could rely \emph{directly} 
on the machinery proposed by \ZpL. At least for now, bringing together 
rapidity and stability still requires a substantial human contribution 
and a careful special study of all parameters.

Nevertheless, we do believe that \ZpL can be extremely helpful to 
whoever wants to design a fast and stable $p$-adic algorihm for a couple 
of reasons. First, it provides solutions 
to automatically detect which steps of a given algorithm are stable and 
which ones are not. This way, it highlights the parts of the algorithm 
on which the researcher has to concentrate its efforts.
Second, recall that a classical strategy to improve stability consists 
in working internally at higher precision. The point is that adjusting 
at best the internal precision (\emph{i.e.} finding the good compromise 
between efficiency and accuracy) is not an easy task in general. It 
turns out that the knowledge of the number of diffused digits of 
precision gives very useful hints in this direction. Indeed, as an 
extreme example, consider the case when there is no diffused digit of 
precision. Then, the optimal precision completely splits over the
variables and there is no need to increase internally the precision. On 
the contrary, when there are many diffused digits of precision, a large 
increment is often required.
Since \ZpL gives a direct access to the number of diffused digits of 
precision, it could be very useful to the designer who is concerned
with the balance between efficiency and accuracy.

\begin{comment}
\subsection{Exploration}
